---
title: "Homework 5 - P8105"
author: "Sergio Ozoria Ramírez (smo2147)"
date: "2025-11-13"
output: github_document
---

# Set up

Potential packages are first loaded to use for our problem sets. Figure preferences were used to set figure sizing and aesthetics for plotting outputs. 

```{r set up, message = FALSE}
library(tidyverse)
library(modelr)
library(rvest)
library(scales)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

# Problem 1

Problem set 1 outlines the function to be used to estimate the probability of sharing a birthday in a calendar year for a group size of 2:50. A figure was plotted to observe these estimates using `ggplot`. 

```{r bday function}
bday_prob = function(n) {
  
  if (!is.numeric(n) || length(n) != 1 || n < 1) {
    
    stop("n must be a single positive number") 
    
    }

  bday_samp = sample(1:365, size = n, replace = TRUE)
  
  any(duplicated(bday_samp))
  
}
```

```{r bday simulation}
set.seed(123)

bday_results =
  tibble(
    n = 2:50,
    bday_probability = map_dbl(
      n, \(x) mean(replicate(10000, bday_prob(x)
                             )
                   )
      )
  )

bday_results
```

```{r figure 1}
bday_results |> 
  ggplot(aes(x = n, y = bday_probability)) +
  geom_hline(yintercept = 0.5, linetype = "dashed") +
  geom_point(color = "blue") +
  scale_y_continuous(labels = percent_format()) +
  labs(
    title = "Figure 1. Probability of Shared Birthday by Group Size",
    x = "Group size (n)",
    y = "Probability of shared birthday"
  )
```

## Figure 1 - Comment

Figure 1 shows that the probability of sharing a birthday increases as the group size increases. For example, in a random group of 23 people, there is a 50% chance that at least two people share the same birthday. We see that the probability of shared birthdays among two people increases closer to 100% with at least 40 people in a random sample. This is because we're making comparisons based on matches, so the rate of possible shared birthdays increases at a faster rate than the group size

# Problem 2

```{r mu & pval function}
sim_test = function(mu, n = 30, sigma = 5) {
  
  x = rnorm(n, mean = mu, sd = sigma)
  
  t_test = t.test(x, mu = 0)
  
  broom::tidy(t_test) |> 
    select(estimate, p.value)
  
}
```


```{r}
sim_mu0 = 
  tibble(
    iter = 1:5000) |> 
  mutate(
    tt_results = map(iter, \(i) sim_test(mu = 0))
  ) |> 
  unnest(tt_results)

sim_mu0
```


```{r}
sim_mu_06 =
  expand_grid(
    mu = 0:6,
    iter = 1:5000
  ) |> 
  mutate(
    tt_results = map(mu, sim_test)
  ) |> 
    unnest(tt_results)

sim_mu_06
```


```{r figure 2}
power_df =
  sim_mu_06 |> 
  group_by(mu) |> 
  summarize(power = mean(p.value < 0.05))

power_df |> 
  ggplot(aes(x = mu, y = power)) +
  geom_line(color = "skyblue") +
  geom_point(color = "blue", size = 2) +
  labs(
    title = "Figure 2. Association between Effect Zize and Power",
    x = "True population mean",
    y = "Power (Prob. of Rejection Null)"
  )
```

## Figure 2 - Comment

Figure 2 shows that the proportion of rejecting the null, meaning power, increases as the true population mean moves further away from 0. Notably, when the true population mean increases from 2 to 3, the proportion of rejecting the null increases significantly, closer to a power of 1

```{r}

```



